---
title: "R Notebook"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---
import the data from the library
```{r}
library("MASS")
head(Cars93)
```


*Histograms* for Cars' price and 93 Models of 1993 Cars
```{r}
hist(Cars93$Price,xlab ="Price ( x $1,000)" , xlim = c(0,70),main = 
"Prices of 93 Models of 1993 Cars",probability = TRUE)
lines(density(Cars93$Price))
```


*table*, illurstrating the types and frequencies of cars. First, make a kind of table which can use 'barplot()'
```{r}
table(Cars93$Type)
```


Then, we can create the graph with above created table.
```{r}
#barplot(table(Cars93$Type),ylim = c(0,25),xlab = "Type",ylab = "Frequency",
 #       axis.lty = "Solid",space = .05)
```


*Pie graphs* , this type of graph couldn't be more strightforward.
```{r}
pie(table(Cars93$Type))
```


*doc charts*, This is one of those infrequent cases where the independent varibale is on the y-axis and the dependent variable is on the x-axis.
let's first create the necessary vectors, chich turn the table into a data frame:
```{r}
type.frame <- data.frame(table(Cars93$Type))
type.frame
```


After you have the data frame, this line produces the doc chart:
```{r}
dotchart(type.frame$Freq,type.frame$Var1)
```


### Bar plots revisted
If the dependent variable is a data point rather than a frequency, this impress us to think about a data visualizing that reveal of those data.
let's first create a vector of the values in the cells:
```{r}
#create a vector of the values in the cells
rev.values <- c(1000,1300,1300,1100,1400,800,1200,1500,
                1850,2330,860,1300,1400,1600,1970,570,380,
                450,465,580,155,190,210,250,300)
```
Then, turns this vector into a matrix.
```{r}
space.rev <- matrix(rev.values,nrow = 5,byrow = T)
```
Finally, supply column names and row names to the matrix:
```{r}
colnames(space.rev)<- c("1990","1991","1992","1993","1994")
rownames(space.rev) <- c("Commercial Satellites Delivered",
                         "Satellite Services","satellite Ground Equipment",
                         "Commercial Launches","Remote Sensing Data")
space.rev
```
create a vector of colors for the bars:
```{r}
color.names = c("black","grey25","grey50","grey75","white")
```
And now for the plot:
```{r}
barplot(space.rev,beside = T,xlab = "Year",ylab = "Revenue (x $1,000)",col = color.names)
#add legend.
legend(1,2300,rownames(space.rev),cex = .7,fill = color.names,bty = "n ")
```




### Scatter plots
*sctter plot*, to visualize the relationship between horsepower and MPG for city driving. with xlab = horsepower and ylab = MPG.City
```{r}
plot(Cars93$Horsepower,Cars93$MPG.City,xlab = "Horsepower",ylab = "MPG.City",main = "MPG City vs
     Horsepower")
```


Alternative way for declaring ,*~* means depends on the following variable.
```{r}
plot(MPG.city ~ Horsepower,data = Cars93,xlab = "Horsepower",ylab = "MPG.City",
     main = "MPG City vs Horsepower")
```


### A plot twist
R enable you to change the symbol that depicts the points in the graph. For default, the symbol is an empty circle. To change the symbol, which is called the plottiung charecter, set the argument *pch*. R has a set of build-in numberical values _(0-25)_ for *pch* that correspond to a set of symbols. The values _0-15_ correspond to unfilled spaces, and _16-25_ are filled.
The default value is _1_. To change the plotting character to squeares, set *pch* to _0_. For **Triangle**, it's _2_, and for filled circles it's _16_:
```{r}
#the charecters are filled circles.
plot(Cars93$Horsepower,Cars93$MPG.city,xlab = "Horsepower",ylab = "MPG City",main =
       "MPG City vs Horsepower",pch = 16)
```
We're not limited to the build-in numberical values for *pch*. Here, for example, is an interesting touch: we can draw each point in the plot as the number of _cylinders_ in the corresponding car, rather than as a symbol.
Need to notice that pass to *pch* is a character rather than a number. Then, to force the Cylinders-value to be character, apply 'as.character()' to _Cars93$Cylinders_:
```{r}
pch.chara = as.character(Cars93$Cylinders)
```
and the 'plot()' function is
```{r}
plot(Cars93$Horsepower,Cars93$MPG.city,xlab = "Horsepower",ylab = "MPG City",main =
       "MPG City vs Horsepower",pch = pch.chara)
```


### Sctter plot matrix
R priovides a nice way to visualizing relationships among more than two variables. The R function for plotting this funciton works with numberical columns from a matrix or a data frame.For convenience, create a data frame that's a subset of the _Car93_ data frame. This data frame consists of just the three variables to plot. The function 'subset()' handles that nicely.
```{r}
cars.subset <- subset(Cars93,select = c(MPG.city,Price,Horsepower))
head(cars.subset)
```
'subset()' has created a vector of exactly what to select out of *Car93*. And now, use 'pair()' function to see the visualizing.
```{r}
pairs(cars.subset)
```


### Gradeating to ggplot2
The Base R graphice toolset will get start, but it doesn't shine at visualization, it's a good idea to learn *ggplot2*. Create by R-megastar Hadley Wickham, the _'gg'_ in the package name stands for "Grammar of graphics" and that's a good indicator of what's ahead.
So a _"grammar of graphics"_ is a set of rules for combining graphics components to produce graphs. Wilkinson proposed that all graphs have underlying common components -- like data, a coordinate system (the x- and y- axis you know so well, for example), statistical transformaitons (like frequency counts), and objects within the graph (e.g., dots, bars, lines, or pie slices), to name a few.

#### Histograms
A graph starts with *ggplot()*, which takes two arguments. The first argument is the source of the data. The second argument maps the data components of interest into components of the graph. The function that doese the job is *aes()*.
To begin a histogram for _price_ in *Cars93*, the function is
```{r}
library("ggplot2")
ggplot(Cars93,aes(x = Price))
```
The *aes()* function associates _Price_ with the x-axis. In ggplot-world, this is called an _aesthetic mapping_. In fact, each arguments to *aes()* is called an _aesthtic_.

And what about that histogram? How do you put it into this blank grid? You have to add something indicating that you want to plot a histogram and let R take care of the rest. What you add is a **geom** function ("Geom" is short for "geometric object").
How do we add **geom_histogram()** to *ggplot()*? With a plus sign:
```{r}
ggplot(Cars93,aes(x = Price)) + geom_histogram()
```
This produces above Figure. The grammar rules tell ggplot2 that when the geometric object is a histogram is a histogram, R does the necessary calculations on the data and produces the appropriate plot.

We can customize about the ggplot. Let's modify the graph. Each bar is called _bin_, and by default, ggplot() uses 30 of them. After ploting the histogram, _ggplot()_ displays an onscreen message that advises experimenting with _binwidth_ (which, unsurprisingly, specifies the width of each bin) to change the graph's appearance. Accordingly, we use _binwidth = 5_ as an argumentin _geom\_histogram()_.
With another function, _labs()_, we modify the labels for the axes and supply a title for the graph.

Altogether now:
```{r}
ggplot(Cars93,aes(x = Price)) + geom_histogram(binwidth = 5,color = "black",fill = "white")+ 
  labs(x = "Price (x $1,000) " ,y = "Frequency" , title = "Prices of 93 Models of 1993 Cars")
```


#### Bar plots
```{r}
ggplot(Cars93,aes(x = Type)) + geom_bar()+ labs(y = "Frequency" , title= "Car Type and Frequency in Cars93")
```


#### Doc charts
Making a doc chart begins much the same as in the base R: create a table for Type, and turn the table into a data frame.
```{r}
type.frame <- data.frame(table(Cars93$Type))
```
To ensure that have meaningful variable names for the aesthetic mapping, apply the *colnames()* function to name the columns in the data frame.
```{r}
colnames(type.frame) <- c("Type","Frequency")
type.frame
```
On the graph. To orient the doc chart, map _Frequency_ variable is on the x-axis and _Type_ to the y-axis:
```{r}
ggplot(type.frame , aes(x = Frequency , y = Type))
```
The _geom_ function for the doc chart is *geom_point()* is available. So this code:
```{r}
ggplot(type.frame, aes(x = Frequency , y = Type)) + geom_point()
```
A couple of modifications are in order. First, with a graph like this, it's a nice touch to rearrange the categories on the y-axis with respect to how they order on what you're measuring on the x-axis. That necessitates a slight change in the aesthetic mapping to the y-axis:
```{r}
ggplot(type.frame,aes(x = Frequency,y = reorder(Type,Frequency))) + geom_point(size = 4)
# Large dots would make the chart look a little nicer
```

Additional function modify the graph's overall appearance. One family of these functions is called _themes_. One member of this family, _theme\_bw()_, removes the gray background. Adding _theme()_ with appropriate a) removes the vertical lines in the grid and b) blackens the horizontal lines and makes them dotted:
```{r}
theme_bw() + 
  theme(panel.grid.major.x = element_blank(), 
        panel.grid.major.y = element_line(color = "black",
                                          linetype = "dotted"))
```


Finally, _lab()_ changes the y-axis label:
```{r}
labs(y = "Type")
```
without the change, the y-axis label would be "reorder(Tpye,Frequency)". Though picturesque, that label makes little sense to the average viewer.
```{r}
ggplot(type.frame,aes(x =Frequency, y = reorder(Type,Frequency)))+
  geom_point(size = 4) + theme_bw()+
    theme(panel.grid.major.x = element_blank(), 
        panel.grid.major.y = element_line(color = "black",
                                          linetype = "dotted")) + labs(y = "Type")
```


#### Bar plots re-revisited
Use _ggplot()_ to create one from _space.rev_, the data set I created from the data.

The function _melt()_ turns wides format into long. Another function, _cast()_, does the reverse. These functions are a huge help because they eliminate the need to go schlepping around in spreadsheets to reshape a data set.
So, with _reshape2_ in the library.

```{r}
library("reshape2")
space.melt <- melt(space.rev)
```
So, that's really all there is to it.
```{r}
# look at those first few rows
head(space.melt)
```
Next, give meaningful names to the columns:
```{r}
colnames(space.melt) <- c("Industry","Year","Revenue")
head(space.melt)
```
And now we are ready to roll. Start with _ggplot()_. The aesthetic mappings are strightforward:
```{r}
ggplot(space.melt,aes(x = Year,y= Revenue,fill = Industry)) + 
  geom_bar(stat = "identity",position = "dodge" , color = "black")
```
```{r}
ggplot(space.melt , aes(x = Year, y = Revenue, fill = Industry)) + geom_bar(stat = "identity",position = "dodge",color = "black")+
  scale_fill_grey(start = 0,end = 1)+
  labs( y = "Revenue (X $1,000)")+
  theme_bw()+
  theme(panel.grid.major.x = element_blank())
```


#### Scatter plots
A scatter plot is a great way to show the relationship between two variables, liks horsepower and miles per gallon for city driving. And _ggplot()_ is a great way to draw the scatter plot. The grammar of this will be easy:
```{r}
ggplot(Cars93,aes (x = Horsepower, y = MPG.city)) + geom_point()
```

Instead, each point is the number of cylinders, which is a label that appears as a text character.
First, we need and additional *aesthetic* mapping in _aes()_. That mapping is **label**, and set it to Cylinders:
```{r}
ggplot(Cars93, aes(x = Horsepower,y = MPG.city , label = Cylinders))+ geom_point()
```


Use the theme functions to make the graph's appearance look like the one showning as before. As in the doc chart example, _theme\_bw()_ eliminates the gray background. The _theme()_ function eliminates the grid:
```{r}
#theme(panle.grid = element_blank())
```
_element\_black()_ is a function that draws a blank element.
Altogether:
```{r}
ggplot(Cars93,aes(x = Horsepower, y = MPG.city, label = Cylinders)) + geom_text() + theme_bw()+
  theme(panel.grid = element_blank())
```



#### Scatter plot matrix
The ggplot2 package had a function called _plotparis()_ that did something similar, but not anymore. GGally, a package build on ggplot2, provides _ggparis()_ to draw scatter plot matries, and it does this in a flaboyant way.

The GGally package isn't on the package tab. We have to selesct Install and type **GGally** in the Install Package dialog box. When it appears on the Packages tab, click the check box next to it.

Or we can type **install.package("_name of the package_")** in the console for installing/collecting this package.
```{r}
install.packages("GGally")
```

```{r}
library("GGally")
ggpairs(cars.subset)
```
As showing, this one's a beauty. The cells along the main diagonal present density plots of the variables.

The three scatter plots are in the cells below the main diagonal. Rather than show the same scatter plots with the axes reversed in the cells above the main diagonal, each above-the-diagonal cell shows a _correlation coefficient_ that summarize the relationship between the cell's row variable and its column variable.

For a real visual treat, add *Cylinders* to *new.cars.subset*, and then apply _ggpairs()_:
```{r}
new.cars.sub <- subset(Cars93,select = c(MPG.city,Price,Horsepower,Cylinders))
ggpairs(new.cars.sub)
```


#### Box plots
Statisticians use box plots to quickly show how groups differ from one another. As in the base R example, the box for Cylinders and Horsepower. This is the replication of the graph in row 3, col 4 or figure shows above.
At this point, we can probably figure out the ggplot() function:
```{r}
ggplot(Cars93,aes(x= Cylinders , y = Horsepower)) + 
  geom_boxplot()
```

Showing all the data points in addition to the boxes, by adding the *geom* function for points
```{r}
ggplot(Cars93,aes(x = Cylinders , y = Horsepower)) + geom_boxplot()+ geom_point()
```

There are mand points overlap, graphics gurus refer to this as *overploting*.
One way to deal with overplotting is to randomly reposition the points so as to reveal them but not change what they represent. This is called _jittering_. And ggplot2 has a _geom_ function for that: *geom_jitter()*. Adding this function to the code:
```{r}
ggplot(Cars93,aes(x = Cylinders , y = Horsepower)) + geom_boxplot()+ geom_point()+
  geom_jitter()
```







