---
title: "Summarizing it All"
output:  html_document
---

## How many ?

We are still using the same dataset. *"MASS"*, remember to include the _libray("MASS")_.
```{r}
library("MASS")
```

_Cars93_ holds data to 27 variables for 93 cars available in 1993.
```{r}
length(Cars93)
```

So *length()* returns the number of variables in the data frame. The function _ncol()_ does the same thing:
```{r}
ncol(Cars93)
```

Find the row of the dataset, *nrow()*
```{r}
nrow(Cars93)
```

To count the number of USA-originated cars, then we state the condition and then add up all the 1s:
```{r}
sum(Cars93$Origin == "USA")
```

To count the number of non-USA cars in the data frame, we can change the condition to "non-USA", of course, or we can use *!=* -- the "not equal" operation:
```{r}
sum(Cars93$Origin != "USA")
```

Addition conditioning, the number of 4-cylinder USA cars:
```{r}
sum(Cars93$Origin == "USA" & Cars93$Cylinders == 4)
```

or, we can also do it (without $-sign):
```{r}
with(Cars93,sum(Origin == "USA" & Cylinders == 4))
```


## The High and the low

Two descriptive statistics that need no introduction are the maximum and minimum value in a set of scores:
```{r}
Horsepower.USA.Four <- with(Cars93, Horsepower[Origin == "USA" & Cylinders == 4])
max(Horsepower.USA.Four)
min(Horsepower.USA.Four)
```

If we happen to need both values at the same time:
```{r}
range(Horsepower.USA.Four)
```


## The Moment

In statistics, moments are quantities that are related to the shape of a set of numbers. By "shape of a set of numbers," I mean "what a histogram based on the number looks like" -- how spread out it is, how symmetric it is, and more.

A _row moment_ of order k is athe average of all numbers in the set, with each number raised to the kth power before we average it.So the *first* raw moment is the arithmetic mean. The *scound* raw moment is the average of the squared scores. The *third* raw moment is the average of the ucbed scores, and so on.

A *central moment* is based on the average of *deviations* of numbers from their mean. If we square the deviations before we average them, we have the *scond* central moment. If we cube the deviations before we averagt them, that's the *third* central moment. raise each one to fourth power before we average them, then we have the *forth* central moment.

### A teachable moment

Here is a function for calculating a central moment of a vector:
```{r}
cen.mom <- function(x,y) {
  mean((x-mean(x))^y)
}
```

Here are the second, third, and fourth central moments:
```{r}
Horsepower.USA <- with(Cars93,Horsepower[Origin == "USA"])
cen.mom(Horsepower.USA,2)
cen.mom(Horsepower.USA,3)
cen.mom(Horsepower.USA,4)
```


## Skewness

For the symmetric histogram, the skewness is **0**. For the histogram -- the one that tails off to the right -- the value of the skewness statistic is *positive*. It's also said to be "skewed to the right" for the third histogram (which tails off to the left), the value of the skewness statistic is *negative*. It's also said to be "skewed to the left".

Now for a formula. let it be $M_k$ represent the kth central moment. To calculate skewness, it's

\[
skewness = \frac{\sum(X-\bar X)^3}{(N-1)s^3}
\]

In English, the skewness of a set of numbers is the third central moment divided by the second central moment raised to the three-halves power. With R function we have defined:
```{r}
cen.mom(Horsepower.USA,3)/cen.mom(Horsepower.USA,2)^1.5
```




































